\Huge Conferences in the Era of Expensive Carbon

### 2019-06-23 main.tex
Complete draft

How should ACM respond to these facts?  Two years ago, SIGPLAN (ACM's
Special Interest Group on Programming Languages) convened an ad hoc Climate
Committee\footnote{https://www.sigplan.org/Resources/Climate/} to research
this question.  After investigating many
options~\cite{ClimateCommitteReport} and discussing intensively with SIGPLAN
members and conference organizers, ACM staff, and other SIG leaders, we have
two concrete proposals. First, all ACM conferences should publicly {\em
  account} for CO2e emitted as a result of putting them on. Second, ACM
should put a \emph{price} on carbon in conference budgets, to create
pressure on organizers to reduce their footprints.

@@

Our first proposal is simple: {\bf Every ACM-sponsored conference should
calculate and publicly report its carbon footprint.} Most conferences'
footprints will be dominated by participants air travel, but the accounting
should go beyond this to include air travel to in-person PC meetings and
estimated hotel and food CO2e footprints.

@@

SIGPLAN has been considering how to reduce CO2e usage over the past couple
of years, informed by an accounting of its CO2e expenditures. This
discussion has led the organizers of two of its flagship conferences (POPL
and ICFP) to switch from in-person to on-line program committee meetings,
joining a trend among other SIGs, most of which now use virtual PC meetings
for some or all of their flagship conferences. It has also prompted
conferences to increase their investments in livestreaming and video
recording to support remote participation.

@@

Beyond encouraging conferences to take these sorts of simple steps, public
accounting of conference emissions footprints will generate valuable data
that future organizers can use to optimize their own footprints.  Emissions
data should be published both in aggregate form (total footprint for the
conference and per-participant footprint), for quick reference, and in a raw
form (a suitably anonymized record for each individual attending each
conference, including originating city), and it should be made available in
a central place and in a uniform format, to facilitate analysis.

ACM can provide tools for gathering and publicizing this data. As a start,
SIGPLAN recently worked with ACM's SIG Governing Board to develop an
air-travel-focused CO2e calculator for conferences. Users can upload
conference registration data, and the calculator will estimate the CO2e cost
of air travel for those sufficiently far away. This calculator could be
expanded to account for other forms of conference emissions.

@@

While publicly accounting for CO2e expenditures will incentivize
some CO2e reductions, it will probably not, by itself, induce major changes
to the conference system. 
%
Science is a fundamentally social process, and the conference system
accelerates scientific research through high-bandwidth interaction,
direct dissemination of results, network building, and serendipitous
cross-fertilization.
%
Organizers and attendees will be reticent to make changes that could
harm the scientific benefits they have all regularly experienced.

@@

The two visualizations of registration data from some recent conferences in
Figure~\ref{fig:confs} illustrate some of the difficulties.  The top diagram
shows an estimated per-participant CO2e footprint for each instance of the
four major SIGPLAN conferences over the past ten years (excluding a few for
which we had trouble getting data), with larger dots representing more
emissions.  Eyeballing this diagram, it appears that carbon-conscious
conference planners should hold all of these conferences either in the
northeast US or in western Europe every year.  But the bottom visualization
tells a different story.  Here, the horizontal colored bar at the bottom
represents the continent on which each conference was held, and each
vertical bar gives a breakdown of the participants in that conference,
colored according to the continent they gave as their home address.  Again,
from a superficial glance at the colors, it is clear that, though a minority
come from far away, the majority of participants in each case are from the
region where the conference is being held---that is, there is a large amount
of local participation.\iflater\bcp{That point could be strengthened by
  adding some numbers (just in the text, not a figure) about what percentage
  of conference attendees attend even one additional instance of the same
  conference...}\fi{} From this, one might conclude that always locating
conferences in the same one or two places would significantly damage the
research community by discouraging participation from other parts of the
world. 
%
%
% \bcp{Explain the two diagrams.  Discuss how they can be used to draw
%   contradictory conclusions.  E.g., one of the many footprint-reduction
%   measures that have been discussed in SIGPLAN~\cite{ClimateCommitteReport}
%   is locating conferences near their historical ``center of mass.''  Looking
%   at the top diagram, one might thus be tempted to locate all conferences in
%   the northeast US or western Europe.  But the bottom figure shows that much
%   of the participation in each instance of a conference is local; always
%   holding a conference at its ``center of mass'' will damage local
%   communities everywhere else.  (Maybe this would be a good place to refer
%   to~\cite{ClimateCommitteReport}---we can enumerate some of the
%   alternatives that one might have hoped the data would help us choose
%   between.)}  
% %
% \bcp{Maybe we can replace the bottom part of the figure by some clearer way
%   of visualizing the observation that much of the attendance at each
%   conference is local?  E.g., what if, instead of color-coding the different
%   sections of the bars by continent, we instead color-coded them by distance
%   traveled (in buckets of 500 miles, say)?}
%
These disparate conclusions suggest that reducing conference emissions will
require genuinely hard painful compromises.  The impulse to avoid or
postpone thinking about it is entirely understandable!

@@

@@

\bcp{I wish we could put a paragraph here about the long-term picture and
  the need to rethink scientific exchange and ACM's business model.  That
  will please our grumpy reviewer, and I think it's actually a point worth
  making.\cite{NCN}}

>>>>

How should ACM respond to these facts?  Two years ago, SIGPLAN (ACM's
Special Interest Group on Programming Languages) convened an ad hoc Climate
Committee\cite{ClimateCommittee} to research this question.  After
investigating many options~\cite{ClimateCommitteReport} and discussing
intensively with SIGPLAN members and conference organizers, ACM staff, and
other SIG leaders, we have two concrete proposals. First, all ACM
conferences should publicly {\em account} for CO2e emitted as a result of
putting them on. Second, ACM should put a \emph{price} on carbon in
conference budgets, to create pressure on organizers to reduce their
footprints.

@@

Our first proposal is simple: {\bf Every ACM-sponsored conference should
publicly report its carbon footprint.} Most conferences'
footprints will be dominated by participants air travel, but the accounting
should go beyond this to include air travel to in-person PC meetings and
estimated hotel and food CO2e footprints.

@@

As a case in point, SIGPLAN has been considering how to reduce CO2e usage
over the past couple of years, informed by an accounting of its own CO2e
expenditures. This discussion has led the organizers of two of its flagship
conferences (POPL and ICFP) to switch from in-person to on-line program
committee meetings, joining a trend among other SIGs, most of which now use
virtual PC meetings for some or all of their flagship conferences. It has
also prompted several conferences to increase their investments in
livestreaming and video recording to support remote participation.

@@

Beyond encouraging conferences to take these sorts of simple steps, public
accounting of conference emissions will generate data that future organizers
can use.  To this end, emissions data should be published both in aggregate
form (total and per-participant footprint), for quick reference, and in raw
form (a suitably anonymized record for each individual attending each
conference, including originating city), and it should be made available in
a central place and in a uniform format, to facilitate analysis.

ACM should provide tools for gathering and publicizing this data. As a
start, SIGPLAN recently worked with ACM's SIG Governing Board to develop an
air-travel-focused CO2e calculator for conferences. Users can upload
conference registration data, and the calculator will estimate the CO2e cost
of air travel for those sufficiently far away. This calculator could be
expanded to account for other forms of conference emissions.

@@

While publicly accounting for CO2e expenditures will incentivize some CO2e
reductions, it is not likely, by itself, to induce major changes.
%
Science is a fundamentally social process, and the conference system
accelerates scientific research through high-bandwidth interaction,
direct dissemination of results, network building, and serendipitous
cross-fertilization.
%
Organizers and attendees will be reticent to make changes that might impact
these benefits.

@@

The two visualizations of registration data from some recent conferences in
Figure~\ref{fig:confs} illustrate this difficulty.  The top diagram shows an
estimated per-participant CO2e footprint for each instance of the four major
SIGPLAN conferences over the past ten years (excluding a few for which we
had trouble getting data), with larger dots representing more emissions.
Eyeballing this diagram, it appears that carbon-conscious conference
planners should hold all of these conferences either in the northeast US or
in western Europe every year.  But the bottom visualization tells a
different story.  Here, the horizontal colored bar at the bottom represents
the continent on which each conference was held, and each vertical bar gives
a breakdown of the participants in that conference, colored according to the
continent they gave as their home address.  Again, from a superficial glance
at the colors, it is clear that, though a minority come from far away, the
majority of participants in each case are from the region where the
conference is being held---that is, there is a large amount of local
participation.\iflater\bcp{That point could be strengthened by adding some
  numbers (just in the text, not a figure) about what percentage of
  conference attendees attend even one additional instance of the same
  conference...}\fi{} From this, one might conclude that always locating
conferences in the same one or two places would significantly damage the
research community by discouraging participation from other parts of the
world---indeed, perhaps conferences should move around as much as
possible. 
%
%
% \bcp{Explain the two diagrams.  Discuss how they can be used to draw
%   contradictory conclusions.  E.g., one of the many footprint-reduction
%   measures that have been discussed in SIGPLAN~\cite{ClimateCommitteReport}
%   is locating conferences near their historical ``center of mass.''  Looking
%   at the top diagram, one might thus be tempted to locate all conferences in
%   the northeast US or western Europe.  But the bottom figure shows that much
%   of the participation in each instance of a conference is local; always
%   holding a conference at its ``center of mass'' will damage local
%   communities everywhere else.  (Maybe this would be a good place to refer
%   to~\cite{ClimateCommitteReport}---we can enumerate some of the
%   alternatives that one might have hoped the data would help us choose
%   between.)}  
% %
% \bcp{Maybe we can replace the bottom part of the figure by some clearer way
%   of visualizing the observation that much of the attendance at each
%   conference is local?  E.g., what if, instead of color-coding the different
%   sections of the bars by continent, we instead color-coded them by distance
%   traveled (in buckets of 500 miles, say)?}
%
These disparate conclusions suggest that reducing conference emissions will
require genuinely painful compromises.  The impulse to avoid thinking about
it is entirely understandable!

@@

One potential sticking point is that some strategies for reducing carbon
footprints---such as encouraging some (or all\cite{NCN}!)  participants to
attend remotely---will also reduce conference revenues; in the limit, this
could significantly impact the income stream of ACM itself, which could make
emissions reduction politically unpalatable unless ACM's conference-focused
business model is also adjusted.  Planning for this eventuality should start
now.

% \SECTION{Prepare for Deeper Changes}

% Our two concrete proposals aim to make conferences much more
% carbon-efficient, allowing ACM to play its part in the 40\% reduction in
% emissions that is needed over the next decade.  But this is just a first
% step.  Looking further ahead, the challenge will be to eliminate {\em all}
% greenhouse emissions by around 2050 and then move into global negative
% emissions during the second half of the century.

% \bcp{I wish we could put a paragraph here about the long-term picture and
%   the need to rethink scientific exchange and ACM's business model.  That
%   will please our grumpy reviewer, and I think it's actually a point worth
%   making.\cite{NCN}}

@@
